This PDF is a Git Repository\\Containing its Own \LaTeX\ Source\\and a Copy of Itself

### 2017-05-08 article.tex
Filled in the section on Delta Objects

%% TODO: Talk about the Delta Ref file format

>>>>

There are two types of delta objects: \emph{references}~(object
type~7) and \emph{offsets}~(object type~6).  Reference delta objects
contain an additional 20~bytes at the end of the header before the
zlib-compressed delta data. These 20~bytes contain the SHA-1 hash of
the base object to which the delta should be applied.  Offset delta
objects are exactly the same, however, instead of referencing the base
object by its SHA-1 hash, it is instead represented by a negative byte
offset to the start of the object within the pack file.  Since a
negative byte offset can typically be encoded in two or three bytes,
it's significantly smaller than a 20-byte SHA-1 hash.  One must
understand how these offset delta objects are encoded if---say, for
some strange, masochistic reason---one wanted to change the order of
objects within a packfile, since doing so would break the negative
offsets. (Foreshadowing!)

One would
\emph{think} that git would use the same multi-byte length encoding that they
used for the uncompressed object length.  But no! This is what we have
to go off of from the git documentation:
\begin{center}
\begingroup
\setbox9=\hbox{\footnotesize\verb|for n >= 2 adding 2^7 + 2^14 + ... + 2^(7*(n-1))|}
\begin{minipage}{\wd9}
\footnotesize\begin{verbatim}
n bytes with MSB set in all but the last one.
The offset is then the number constructed by
concatenating the lower 7 bit of each byte, and
for n >= 2 adding 2^7 + 2^14 + ... + 2^(7*(n-1))
to the result.
\end{verbatim}
\end{minipage}
\endgroup
\end{center}
Right. Some experimenting resulted in the following decoding logic
that appears to work:
\begin{center}
\begingroup
\setbox9=\hbox{\footnotesize\verb|        reference += (1 << (7 * (bytes_read - 1)))|}
\begin{minipage}{\wd9}
\footnotesize
{\color{blue}\verb|def|}\verb| decode_obj_ref|{\color{gray}\verb|(|}\verb|data|{\color{gray}\verb|):|}\\
\verb|    bytes_read = 0|\\
\verb|    reference = 0|\\
\verb|    |{\color{blue}\verb|for|}\verb| c |{\color{blue}\verb|in map|}{\color{gray}\verb|(|}\verb|ord, data|{\color{gray}\verb|):|}\\
\verb|        bytes_read += 1|\\
\verb|        reference <<= 7|\\
\verb|        reference += c & 0b01111111|\\
\verb|        |{\color{blue}\verb|if not|}\verb| |{\color{gray}\verb|(|}\verb|c & 0b10000000|{\color{gray}\verb|):|}\\
{\color{blue}\verb|            break|}\\
{\color{blue}\verb|    if|}\verb| bytes_read >= 2|{\color{gray}\verb|:|}\\
\verb|        reference += |{\color{gray}\verb|(|}\verb|1 << |{\color{gray}\verb|(|}\verb|7 * |{\color{gray}\verb|(|}\verb|bytes_read - 1|{\color{gray}\verb|)))|}\\
\verb|    return reference, bytes_read|\\
\end{minipage}
\endgroup
\end{center}

The rabbit hole is deeper still; we haven't yet discovered the content
of the compressed delta objects, let alone how they are applied to
base objects.  At this point, we have more than sufficient knowledge
to proceed with the PoC, and my canary died ages ago. Aditya Mukerjee
did a good job of explaining the process of applying deltas in his
blog post, so I will stop here and proceed with the polyglot.
